Hydrodynamics and rheology of weakly adhesive vesicle doublets

### 2019-01-28 vesicle_code/docs/adhesion/adhesion.tex
some minor changes by yny

The dynamics of weakly adhesive vesicles under various flowing conditions is investigated numerically using a boundary integral formulation with
adaptive time-marching that facilitates efficient and accurate computations of vesicle hydrodynamics under adhesion. In a quiescent flow, a lubrication analysis illustrates the scaling in 
the draining dynamics with respect to the adhesion strength and length scale. In a microfluid trap, results from boundary integral simulations show that there exists a critical extensional flow rate above which the weak adhesive interaction is overcome by the hydrodynamic diverging stream, thus providing a simple method to measure the adhesion strength between two vesicle membranes. In a planar shear flow, numerical simulations reveal that a vesicle doublet may form as they approach each other in a planar shear flow provided that the adhesion
strength is sufficient large at a given vesicle reduced area. Once a doublet the effective shear viscosity of a dilute suspension of vesicle doublets is found 
to be a function of both reduced area and the adhesion strength.
Results from these numerical studies and analysis shed light on the hydrodynamic effects of adhesive interactions between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

%
% for example, 
%it was shown recently that vesicles may align to form chains along the direction of the externally applied DC electric field with ensuing consequences to suspension rheology \cite{wu2018pairwise}.
%
Collective hydrodynamics of vesicles naturally is dictated by the individual
vesicle-vesicle interactions.  
In a quiescent flow,
vesicle-vesicle adhesion leads to formation of vesicle doublets \cite{ZiherlSvetina2007_PNAS} or clusters \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}.
As a model for red blood cell (RBC) aggregates, an idealized model for adhesive vesicle-vesicle interactions gives rise to 
vesicle configurations that are observed in experiments of fibrinogen-induced RBC aggregates \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}. 
Using the L-J potential for pairwise point-point interaction between two vesicle membranes without the detailed molecular details of
adhesive interactions between RBCs, Flormann {\it et al.} found the contact zone to buckle under strong adhesion \cite{FlormannAouane2017_SciReports}. 
A direct consequence of the buckling is the
sigmoidal shape of the adhesion region, which is also observed in RBC doublets and explained in a slightly different model \cite{ZiherlSvetina2007_PNAS}. 
It remains unknown how adhesive interactions between vesicles might lead to different vesicle hydrodynamics that results in
different rheological properties of a vesicle suspension. Such studies will provide useful insight to the rheology of blood flow and how to use nano solutes to control the hypertension
by adjusting the adhesive interactions between RBCs.
%

In numerical simulations of  a vesicle suspension, the adhesive interaction between two vesicles is often ignored for numerical convenience
because a smaller time step is often needed to resolve the dynamics of the lubrication thin film between two membranes with adhesion than without.
However, adhesive interactions  between RBCs lead to different RBC clusters that are expected to alter the rheological properties of the RBC suspension \cite{NeuMeiselman2002_BiophysJ,SvetinaZiherl2008_Bioelectrochemistry}.
Hydrodynamic simulations of adhesive vesicles in a quiescent flow have revealed physical insights to the equilibrium shapes of RBC aggregates in both experiments 
\cite{FlormannAouane2017_SciReports} and theory \cite{ZiherlSvetina2007_PNAS}. The main goal in this work is to investigate the hydrodynamics of adhesive vesicles in flowing
conditions common in microfluidic vesicle flow, such as a planar shear flow or an extensional flow.

%
The electrostatic
natures of lipid molecules (a hydrophobic tail and a hydrophilic head
with an electric dipole) complicate the interactions between a lipid
bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces,PerutkovaFrank-Bertoncelij2013_CSB} 
or a solid (such as a glass substrate
or a nano particle).  It is reported that such complicated electrostatic contribution in the adhesive interactions between two membranes can be well approximated by the 
L-J-type potential \cite{FlormannAouane2017_SciReports}, which consists of a long-range attraction component and a
short-range repulsion component~\cite{Book_IntermolecularSurfaceForces},
and the combination of the two gives rise to adhesion at the separation
distance that minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. Such adhesive
potentials are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is essential for membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.

>>>>

The dynamics of weakly adhesive vesicles under various flowing conditions is investigated numerically using a boundary integral formulation with
adaptive time-marching that facilitates efficient and accurate computations of vesicle hydrodynamics under adhesion. In a quiescent flow, two nearby vesicles 
move slowly towards each other under the adhesive potential, pushing out fluid in the gap to  
form a doublet at equilibrium. A lubrication analysis on such draining of the
thin film gives the dependencies of draining time on adhesion strength and separation distance that are in good agreement with results from
boundary integral simulations.
%
In a microfluid trap where the stagnation of an extensional flow  is placed in the middle of a vesicle doublet through an active control loop,
novel dynamics of a vesicle double tis observed. Results from boundary integral simulations show that there exists a critical extensional flow rate above which the weak adhesive interaction is overcome by the hydrodynamic diverging stream, thus providing a simple method to measure the adhesion strength between two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle doublet may form as they approach each other in a planar shear flow provided that the adhesion
strength is sufficient large at a given vesicle reduced area. Once a doublet the effective shear viscosity of a dilute suspension of vesicle doublets is found 
to be a function of both reduced area and the adhesion strength.
Results from these numerical studies and analysis shed light on the hydrodynamic effects of adhesive interactions between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

%
% for example, 
%it was shown recently that vesicles may align to form chains along the direction of the externally applied DC electric field with ensuing consequences to suspension rheology \cite{wu2018pairwise}.
%
Collective hydrodynamics of vesicles is dictated by the individual
vesicle-vesicle interactions.  
In a quiescent flow,
vesicle-vesicle adhesion leads to formation of vesicle doublets \cite{ZiherlSvetina2007_PNAS} or clusters \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}.
As a model for red blood cell (RBC) aggregates, an idealized model for adhesive vesicle-vesicle interactions gives rise to 
vesicle configurations similar to those observed in experiments of fibrinogen-induced RBC aggregates \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}. 
Using the L-J potential for point-point interaction between two vesicle membranes without the detailed molecular details of
adhesive interactions between RBCs, Flormann {\it et al.} found the contact zone to buckle under strong adhesion \cite{FlormannAouane2017_SciReports}. 
A direct consequence of the buckling is the
sigmoidal shape of the adhesion region, which is also observed in RBC doublets and explained in a slightly different model \cite{ZiherlSvetina2007_PNAS}. 
It remains unknown how adhesive interactions between vesicles might lead to different vesicle hydrodynamics that results in
different rheological properties of a vesicle suspension. Such studies will provide useful insight to the rheology of blood flow and how to use nano solutes to control the hypertension
by adjusting the adhesive interactions between RBCs.
%

In numerical simulations of  a vesicle suspension, the adhesive interaction between two vesicles is often ignored \cite{Veerapaneni2009_JCP,RahimianVeerapaneniBiros2010_JCP} 
for numerical convenience
because a smaller time step is often needed to resolve the dynamics of the lubrication thin film between two membranes with adhesion than without.
However, adhesive interactions  between RBCs lead to different RBC clusters that are expected to alter the rheological properties of the RBC suspension \cite{NeuMeiselman2002_BiophysJ,SvetinaZiherl2008_Bioelectrochemistry}.
Hydrodynamic simulations of adhesive vesicles in a quiescent flow have revealed physical insights to the equilibrium shapes of RBC aggregates in both experiments 
\cite{FlormannAouane2017_SciReports} and theory \cite{ZiherlSvetina2007_PNAS}. The main goal in this work is to investigate the hydrodynamics of adhesive vesicles in flowing
conditions common in microfluidic vesicle flow, such as a planar shear flow or an extensional flow.

%
The electrostatic
nature of lipid molecules (a hydrophobic tail and a hydrophilic head
with an electric dipole) complicates the interactions between a lipid
bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces,PerutkovaFrank-Bertoncelij2013_CSB} 
or a solid (such as a glass substrate
or a nano particle).  In the absence of an external electric field and ions in the solvents, 
it is reported that 
%such complicated electrostatic component in 
the adhesive interactions between two membranes can be well approximated by the 
L-J-type potential \cite{FlormannAouane2017_SciReports}, which consists of a long-range attraction component and a
short-range repulsion component~\cite{Book_IntermolecularSurfaceForces},
and the combination of the two gives rise to adhesion at the separation
distance that minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. Such adhesive
potentials are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is essential for membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.
